---
title: "Analysis of characteristics in victims of crimes in Toronto"
subtitle: "My subtitle if needed"
author: 
  - Xiyan Chen
thanks: "Code and data are available at: https://github.com/siaxiyann/crime.git"
date: today
date-format: long
abstract: This criterion is linked to a Learning OutcomeAbstract
An abstract is included and appropriately pitched to a non-specialist audience. The abstract answers: 1) what was done, 2) what was found, and 3) why this matters (all at a high level). Likely four sentences. Abstract must make clear what we learn about the world because of this paper. "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(opendatatoronto)
library(ggplot2)
library(dplyr)
```


# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @citeRohan

The remainder of this paper is structured as follows. @sec-data....
The introduction is self-contained and tells a reader everything they need to know including: 1) broader context to motivate; 2) some detail about what the paper is about; 3) a clear gap that needs to be filled; 4) what was done; 5) what was found; 6) why it is important; 7) the structure of the paper. A reader should be able to read only the introduction and know what was done, why, and what was found. Likely 3 or 4 paragraphs, or 10 per cent of total.


# Data {#sec-data}

Some of our data is of penguins (@zaizheli fig-bills), from @opendatatoronto.
A sense of the dataset should be communicated to the reader. The broader context of the dataset should be discussed. All variables should be thoroughly examined and explained. Explain if there were similar datasets that could have been used and why they were not. If variables were constructed then this should be mentioned, and high-level cleaning aspects of note should be mentioned, but this section should focus on the destination, not the journey. It is important to understand what the variables look like by including graphs, and possibly tables, of all observations, along with discussion of those graphs and the other features of these data. Summary statistics should also be included, and well as any relationships between the variables. If this becomes too detailed, then appendices could be used.
This criterion is linked to a Learning OutcomeMeasurement
A thorough discussion of measurement, relating to the dataset, is provided in the data section.
```{r}
analysis_data <- read_csv("../data/raw_data/analysis_data.csv")
```

```{r}
#| label: fig-general_overview
#| fig-cap: Distribution of crime types by year
#| echo: false

ggplot(analysis_data, aes(x = REPORT_YEAR, weight = COUNT_, fill = SUBTYPE)) +
  geom_bar() + labs(title = 'crime cases overview from 2014-2023', y = 'total counts', x = 'report year') + 
  scale_x_continuous(breaks = seq(min(analysis_data$REPORT_YEAR), max(analysis_data$REPORT_YEAR), by = 1))
  #scale_fill_manual(values = c("darkorange","purple","cyan4"),
  #                  guide = "none") +
  #theme_minimal() +
  #facet_wrap(~species, ncol = 1) +
  #coord_flip()
```

The graph @fig-general_overview shows that..

```{r}
#| label: fig-crimetype
#| fig-cap: examing the number of crime case in 4 types from 2014-2024
#| echo: false

ggplot(analysis_data, aes(x = SUBTYPE)) +
  geom_bar() +
  labs(title = 'the number of cases in 4 crime types', y = 'total counts', x = 'crimes type') 
```

in the figure @fig-crimetype it shows

```{r}
#| label: fig-crime
#| fig-cap: examing the number of crime case in 4 types from 2014-2024
#| echo: false

assaulted <- analysis_data %>% 
  filter(SUBTYPE == "Assault") %>%
  group_by(SEX) %>%
  summarise(total_count = sum(COUNT_))

ggplot(assaulted, aes(x = SEX, y = total_count, fill= SEX)) +
  geom_bar(stat = "identity") +
  labs(title = 'Gender Distribution in Assault Cases', y = 'total assault cases', x = 'gender') 
```


Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```


\newpage


# References


